Does pre-crastination explain why some observers are sub-optimal in a visual search task?

### 2023-09-07 RegReport/paper.tex
Update paper.tex

In addition to the registered measures, for efficiency we collected a range of other demographic characteristics and free-report measures from our participants. These are outside the scope of the pre-registration we have presented in this paper, so a full exploration of these additional variables and their relationship to visual search and precrastination will be presented elsewhere. In the interest of open science, however, and to avoid possible confusion by separating or duplicating data, the full set of data analysed in this pre-registration is presented along with these additional variables here: [link to data].

\section{General Discussion}
Although the smaller pilot sample (N=30) suggested participants who pre-crastinate tended to use their peripheral vision to search more efficiently, the larger sample (N=201) did not support that conclusion. While this is a null finding, it is nonetheless meaningful because we had a clear prediction based on data and theory and designed a robust experiment to test it, which underwent rigourous peer review. Both the pilot and the final experiment replicated the previously-established patterns in finding (1) a good split of people who consistently pre-crastinated and people who consistently did not and (2) a broad range of individual differences in visual search efficiency. If a stable relationship between these factors exists and is large enough to build any theories around, this experiment would have detected it. The fact that it did not detect it is useful in ruling out explanations for either behaviour (precrastination or visual search inefficiency) that is grounded in a shared personality characteristic like conscientiousness or impulsivity.

The relationship between conscientiousness and pre-crastination observed previously \cite{sauerberger2019} was not observed here. This could be because the original effect was a statistical accident, or because the effect is small and not present in our sample. One point to note is that our sample was skewed towards the higher end of the conscientiousness scale, likely because unconscientious people are less inclined to participate in psychology experiments. It is possible that the sample in \cite{sauerberger2019} was more representative of the full range of conscientiousness in the population, and was therefore more sensitive to the relationship with precrastination. On the other hand, one might expect our more conscientious sample would have higher precrastination rates than previously observed, and this was not the case. Another interesting and counterintuitive observation from \cite{sauerberger2019} was that procrastination was unrelated to precrastination, and the current results show a small negative correlation, as one would expect if starting tasks too soon is predictive of not starting tasks too late.

Our results provide a relatively definitive answer to a set of similar questions that have been posed with some frequency by other researchers in review comments and in discussions at conferences and research meetings: can we explain variation in individual strategies in these two tasks using these well-established questionnaires? Although the simple answer is "no", the hypothesis that personality traits can explain or predict behaviour in these tasks cannot be completely put to rest by our results, because it is possible the relationships are nonlinear, or there are heterogeneous influences that can drive two people to exhibit the same behaviour, but for different reasons. Nonetheless, had we found the predicted relationships in the larger sample, this would have been an important finding, that would have indicated we can predict how people behave in these tasks based on personality traits (and vice-versa). It was therefore an important possibility to test, even though the results did not bear out the prediction. Moreover, through this project, we have produced a large and very rich dataset with many other measures included, in addition to those reported here. We invite others to use it in hypothesis-generating exploratory analyses of their choosing. We request that anyone using the data in this repository reference this paper.

>>>>

In addition to the registered measures, for efficiency we collected a range of other demographic characteristics and free-report measures from our participants. These are outside the scope of the questions we have addressed in this paper, so a full exploration of these additional variables and their relationship to visual search and precrastination will be presented elsewhere. In the interest of open science, however, and to avoid possible confusion by separating or duplicating data, the full set of data analysed in this pre-registration is presented along with these additional variables here: [link to data].

\section{General Discussion}
Although the pilot data suggested participants who pre-crastinate tended to use their peripheral vision to search more efficiently, the registered study with a much larger sampledid not support that conclusion. While this is a null finding, it is nonetheless meaningful because we had a clear prediction based on data and theory and designed a robust experiment to test it, which underwent rigourous peer review. Both the pilot and the final experiment replicated the previously-established patterns in finding (1) a good split of people who consistently pre-crastinated and people who consistently did not and (2) a broad range of individual differences in visual search efficiency. If a stable relationship between these factors exists and is large enough to build any theories around, this experiment would have detected it. The fact that it did not detect it is useful in ruling out explanations for either behaviour (precrastination or visual search inefficiency) that is grounded in a shared personality characteristic like conscientiousness or impulsivity.

The relationship between conscientiousness and pre-crastination observed previously \cite{sauerberger2019} was not observed here. This could be because the original effect was a statistical accident, or because the effect is not present in our sample. One point to note is that our sample was skewed towards the higher end of the conscientiousness scale, likely because unconscientious people are less inclined to participate in psychology experiments. It is possible that the sample in \cite{sauerberger2019} was more representative of the full range of conscientiousness in the population, and was therefore more sensitive to the relationship with precrastination. On the other hand, one might expect our more conscientious sample would have higher precrastination rates than previously observed, and this was not the case. Another interesting and counterintuitive observation from \cite{sauerberger2019} was that procrastination was unrelated to precrastination, and the current results show a small negative correlation, as one would expect if starting tasks too soon is predictive of not starting tasks too late. Given the inconsistency of this finding between the present results and \cite{sauerberger2019}, however, both of which are samples of over 300 participants, the relationship between these two measures is not strong.

Our results provide a relatively definitive answer to a set of related questions: can we explain variation in individual strategies in precrastination and visual search effiency using these well-established personality questionnaires? Although the simple answer is "no" (or at least, not much), the hypothesis that personality traits can explain or predict behaviour in these tasks cannot be completely put to rest by our results, because it is possible the relationships are nonlinear, or there are heterogeneous influences that can drive two people to exhibit the same behaviour, but for different reasons. Nonetheless, had we found the predicted relationships in the larger sample, this would have been an important finding, that would have indicated we can predict how people behave in these tasks based on personality traits (and vice-versa). It was therefore an important possibility to test, even though the results did not bear out the prediction. Moreover, through this project, we have produced a large and very rich dataset with many other measures included, in addition to those reported here. We invite others to use it in hypothesis-generating exploratory analyses of their choosing. We request that anyone using the data in this repository reference this paper.
